Thesis: draft

### 2021-04-17 draft.tex
plus minus finished stubborn people section in results

\subsection{Including a stubbornness parameter}
\label{stubbpar}
Here, we review the results of introducing stubborn people by including a stubbornness parameter $r_i$. We look at two slightly different case
\begin{itemize}
	\item Making a fraction of the nodes (going from none of the nodes to all the nodes) completely stubborn ($r_i = 1$),
	\item Making all nodes partly stubborn (going from not stubborn at all, $r = 0$ to completely stubborn, $r = 1$).
\end{itemize}
For both cases we will look at both the evolution of the opinions and the formation of echo chambers. \\
Once again, the opinion dynamics model described in Chapter \ref{chap4}, Section \ref{opinionmodel} is implemented with an activation probability equal to $p_{\text{act}} = 0.1$ and a maximum length of the actual time-line $S = 20$.

\newpage
\noindent
\textbf{Making a fraction of the nodes completely stubborn}\\
\newline
We start by making a fraction of the nodes (fracRes) completely stubborn ($r_i = 1$). We go from no stubborn nodes, fracRes $=0$, to all the nodes are stubborn, fracRes $=1$. In all cases, the stubborn nodes are randomly distributed over the network.\\
We will first discuss the formation of echo chambers for an initial $50/50$ opinion distribution, after which the results for the opinion evolution will be shared.

@@

@@

If the fraction of completely stubborn nodes becomes bigger than fracRes $\geqslant 0.25$, the formation of echo chambers starts to decrease. This is probably due to the fact that too many nodes become stubborn, so there are more and more nodes that won't change opinions. This in turn increases the chance of having stubborn neighbors, each with different opinions which disables the formation of echo chambers.

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_frac_compl_stubb_REF_8x7.png}
    \caption{REF}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_frac_compl_stubb_REC_8x7.png}
    \caption{REC}
    \label{REC_frac_compl_stubb}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Formation of echo chambers versus fraction of completely stubborn nodes for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution. \textbf{(a)} Using the REF filtering algorithm; \textbf{(b)} Using the REC filtering algorithm. The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1. This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure). Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$. Each data point is the result of $10^2$ independent simulations.}
\label{echo_vs_frac_complStubb_REF-REC}
\end{figure}
\noindent
So far the results for the formation of echo chambers in the system for an initial $50/50$ opinion distribution and introducing a fraction of completely stubborn agents. But what about the evolution of the opinions over time? Are these affected by introducing stubborn people in the system? \\
\newline
In Figure \ref{ev_op_50_50_stubb_REC} the evolution of the fraction of nodes with opinion 0 for an initial $50/50$ opinion distribution is shown. Data points are shown together with the standard deviation $\sigma$. This figure shows the results for the REC filtering algorithm. The four network models (ER, WS, SBM (high and low modularity) and SBM-WS) are compared to each other. For each model we look at three different fractions of completely stubborn people: no stubborn people, fracRes $=0$; a fraction of completely stubborn people equal to fracRes $=0.4$ and a fraction of completely stubborn people equal to fracRes $=0.8$. Introducing a fraction of stubborn people is not able to break the initial $50/50$ opinion distribution for neither of the network models.\\
\newline
Figure \ref{ev_op_50_50_stubb_PR} shows the results for the evolution of the fraction of opinion 0 in the system for an initial $50/50$ opinion distribution and when using the PR filtering algorithm. Data points are again represented together with the standard deviation $\sigma$. Three different fractions of stubborn people (fracRes $=0$, fracRes $=0.4$ and fracRes $=0.8$) are compared to each other for the four different network models. The results don't differ from those obtained for the REC filtering algorithm. The status-quo is maintained, the initial $50/50$ opinion distribution does not change over time when introducing stubborn people and using the PR filtering algorithm.

@@

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_PR_8x7.png}
    \caption{ER}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_PR_8x7.png}
    \caption{WS}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_PR_8x7.png}
    \caption{SBM}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_PR_8x7.png}
    \caption{SBM-WS}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$). Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$. Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
\label{ev_op_50_50_stubb_PR}
\end{figure}
\noindent
Up till now, we have discussed the results for the formation of echo chambers and the evolution of the opinions when introducing a fraction of completely stubborn nodes in the system for an initial $50/50$ opinion distribution. Lets now look at the results for introducing a fraction of stubborn agents when we are dealing with an unbalanced initial opinion distribution, eg. an initial $20/80$ distribution.\\
\newline

@@

Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80} shows the formation of echo chambers versus the fraction of completely stubborn nodes ($r_i = 1$), fracRes, for an initial $20/80$ opinion distribution (opinion 0 is the minority opinion, opinion 1 is the majority opinion). Figure \ref{PR_frac_compl_stubb_20-80} shows the results for the PR filtering algorithm and Figure \ref{REC_frac_compl_stubb_20-80} shows those for the REC filtering algorithm. The data points are shown together with their standard error $SE$, note that these are very small and almost not visible in the figure. The interpretation of the $y$-axis is slightly different than before. In previous figures, the echo chamber size was interpreted as the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1. Thus we first normalized the echo chambers around opinion 0 and those around opinion 1 and than took the average of both. In case of a $20/80$ distribution we have a very skewed echo chamber distribution, as can be seen in Figure \ref{echo_20_80}, with normalized values going to infinity. We still want to have some average of the echo chambers around opinion 0 and those around opinion 1 to quantify the total formation of echo chambers in the system, but we want to avoid problems with those infinities. Hence, we chose for a slightly different average: first we calculate the total number of like-minded neighborhoods at $t=500$ (thus the sum of the number of echo chambers around opinion 0 and the number of echo chambers around opinion 1) and then we divide that quantity by the total number of like-minded neighborhoods at $t=0$ (hence, we first take the average and than normalize it, instead of normalizing first and then taking the average, as was done before). Note that by taking the average, we neglect the skewed echo chamber distribution. This is, however, not really a problem, since we just want a sense of the total number of echo chambers in the system and how this total number is impacted by the presence of stubborn people and are less interested in the distribution of those echo chambers. \\

@@

When using the PR filtering algorithm, see Figure \ref{PR_frac_compl_stubb_20-80}, we can observe a decrease in the formation of echo chambers with an increasing fraction of stubborn nodes. This behavior is similar to the case of an initial $50/50$ opinion distribution in combination with the PR filtering algorithm as depicted in Figure \ref{echo_vs_frac_complStubb}. Note that for an initial $20/80$ distribution, in combination with the PR filtering algorithm, the formation of echo chambers in the Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity is non-zero (for a fraction of completely stubborn nodes equal to fracRes $=0$), as can also be observed in Figure \ref{echo_20_80}. This behavior is a consequence of the fact that for an initial $20/80$ opinion distribution, in combination with the PR filtering algorithm, the prevalence of the majority opinion strongly increases, as was already explained for Figure \ref{echo_20_80}. The formation of echo chambers for networks with a high clustering and/or a high modularity is slightly higher. It is however not recommended to compare the formation of echo chambers between the different network models when depicting the results as the average of the echo chambers around opinion 0 and those around opinion 1 in case of an initial $20/80$ opinion distribution. The reason for this is that this average neglects the skewed echo chamber distributions as depicted in Figure \ref{echo_20_80}. When comparing the different network models, these skewed distributions should not be overlooked. The goal of this figure is to show the dependency of the formation of echo chambers in the network models on the presence of a fraction of stubborn people and not to compare the different models against each other. \\

@@

So much for the formation of echo chambers in case of an unbalanced initial opinion distribution and introducing a fraction of stubborn people in the system. Lets now discuss the evolution of the opinions in the system. In Figure \ref{ev_op_20_80}, we already saw that for a system with no stubborn agents the REC method does not break the prevalence of the initial opinions in case of an initial $20/80$ opinion distribution. The PR filtering algorithm, on the other hand, resulted in a strong decrease of the prevalence of the minority opinion. Are these results affected by introducing stubborn agents?\\

@@

Figure \ref{ev_op_20_80_stubb_REC} shows the evolution of the fraction of nodes with opinion 0 over time for an initial $20/80$ distribution and when using the REC filtering algorithm. Data point are shown together with their standard deviation $\sigma$. Four network models are compared (ER, WS, SBM (high and low modularity) and SBM-WS). For each network model we compare three different fractions of completely stubborn nodes (fracRes $=0$, fracRes $=0.4$ and fracRes $=0.8$). From this figure it becomes clear that introducing stubborn people in combination with the REC filtering algorithm is not able to break the prevalence of the initial opinions in case of an initial $20/80$ opinion distribution. None of the network topologies, nor the stubborn agents can break the status-quo. This result is similar to the case of an initial $50/50$ opinion distribution in combination with a fraction of stubborn people, as discussed for Figure \ref{ev_op_50_50_stubb_REC}.\\

@@

Figure \ref{ev_op_20_80_stubb_PR} shows the results for the fraction of opinion 0 versus time in case of an unbalanced initial opinion distribution and when using the PR filtering algorithm. Data points are presented along with their standard deviation $\sigma$. Again, we compare four different network models (ER, WS, SBM (high and low modularity) and SBM-WS) and for each model we compare the results for three different fractions of stubborn people (fracRes $=0$, fracRes $=0.4$ and fracRes $=0.8$). If there are no stubborn people, the PR leads to a significant decrease in the prevalence of the minority opinion in case of an initial unbalanced opinion distribution. Introducing a fraction of stubborn agents changes this scenario. For an increasing fraction of stubborn people, the decrease of the minority opinion becomes less and less strong. This result is the same for all the different network topologies. If there are no stubborn people, the PR method forces nodes with the minority opinion to switch to the majority opinion. If we however introduce a fraction of stubborn people in the system, then we will also end up with some stubborn nodes that carry the minority opinion. These stubborn nodes with the minority opinion will not switch to the majority opinion, no matter how hard the PR filtering algorithm pushes them. The consequence is that these stubborn people with the minority opinion hamper the decrease of that minority opinion.

@@

  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_PR_20-80_8x7.png}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_PR_20-80_8x7.png}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_PR_20-80_8x7.png}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_PR_20-80_8x7.png}

>>>>

\subsection{Including a stubbornness parameter}
\label{stubbpar}
Here, we review the results of introducing stubborn people by including a stubbornness parameter $r_i$. We look at two slightly different case
\begin{itemize}
	\item Making a fraction of the nodes (going from none of the nodes to all the nodes) completely stubborn ($r_i = 1$),
	\item Making all nodes partly stubborn (going from not stubborn at all, $r = 0$ to completely stubborn, $r = 1$).
\end{itemize}
For both cases we will look at both the evolution of the opinions and the formation of echo chambers and we will do this for two different initial conditions: a balanced initial opinion distribution ($50/50$) and an unbalanced initial opinion distribution ($20/80$). \\
Once again, the opinion dynamics model described in Chapter \ref{chap4}, Section \ref{opinionmodel} is implemented with an activation probability equal to $p_{\text{act}} = 0.1$ and a maximum length of the actual time-line $S = 20$.

\newpage
\noindent
\textbf{Making a fraction of the nodes completely stubborn}\\
\newline
We start by making a fraction of the nodes (fracRes) completely stubborn ($r_i = 1$). We go from no stubborn nodes, fracRes $=0$, to all the nodes are stubborn, fracRes $=1$. In all cases, the stubborn nodes are randomly distributed over the network.\\
We will first discuss the formation of echo chambers for an initial $50/50$ opinion distribution, after which the results for the opinion evolution will be shared. Then we will discuss the formation of echo chambers and the evolution of the opinions for an initial $20/80$ opinion distribution.

@@

  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_frac_compl_stubb_REF_8x7.png}
  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_frac_compl_stubb_REC_8x7.png}
    \label{REC_frac_compl_stubb}
  \caption[Formation of echo chambers versus fraction of completely stubborn nodes for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution. \textbf{(a)} Using the REF filtering algorithm; \textbf{(b)} Using the REC filtering algorithm. The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1. This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure). Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$. Each data point is the result of $10^2$ independent simulations.}
\label{echo_vs_frac_complStubb_REF-REC}

@@

If the fraction of completely stubborn nodes becomes bigger than fracRes $\geqslant 0.25$, the formation of echo chambers starts to decrease. This is probably due to the fact that too many nodes become stubborn, so there are more and more nodes that won't change opinions. This in turn increases the chance of having stubborn neighbors, each with different opinions which disables the formation of echo chambers.\\
This discrepancy between the results for the PR method and those for the REF/REC method in combination with a small fraction of stubborn nodes can be potentially explained with the help of the following simple example. Assume we have three nodes that form a triangle. Two nodes initially carry opinion 0 and one node carries opinion 1.\\
Lets first discuss the PR method and assume that none of the nodes are stubborn. In case of the PR filtering algorithm this triangle has a high chance of forming an echo chamber (which is obtained if all the nodes adopt the same opinion). This can be easily seen: the one node with opinion 1 is likely to switch to opinion 0, since all his friends carry this opinion and hence, this opinion will be strongly pushed on his time-line (the fact that the PR method wants to promote posts with opinion 1 does not matter here, since there are no neighbors with opinion 1). The two nodes with opinion 0 will have more chance of keeping that opinion than switching to opinion 1 in case of the PR method. This is because the PR method will promote the posts of the friend with the same opinion 0 over the posts of the friend with opinion 1. The end result will be that there is a high chance that all the three nodes will adopt opinion 0 and hence, form an echo chamber. Now assume that we make the node with opinion 1 stubborn. This node is then no longer capable of switching to opinion 0. However, the two nodes with opinion 0 still have more chance of keeping this opinion in case of the PR filtering algorithm, because this filtering algorithm keeps on promoting the posts of the neighbor with the same opinion 0, and hence reduces the visibility of the posts of the stubborn friend with opinion 1. The system will now not form an echo chamber, since the stubborn node with opinion 1 cannot switch to opinion 0 and the PR method reduces the chance of changing the opinion of the nodes with opinion 0 to opinion 1. The conclusion is that, in case of the PR filtering algorithm, making a fraction of the nodes stubborn reduces the ability of the system to form echo chambers compared to the case where there are no stubborn people.\\
What happens if we use the REF/REC filtering algorithm? Again we have one node with opinion 1 and two nodes with opinion 0 in the triangle. Assume first that none of the nodes are stubborn. For the PR method, we argued that in this case, the triangle has a high chance of evolving towards an echo chamber. For the REF/REC method this chance is however smaller, which can be seen as follows. The node with opinion 1 still has a high chance of switching to opinion 0, because all his friends (which are the two other nodes in the triangle) carry this opinion. However, the two nodes with opinion 0 will now not have a higher chance of keeping this opinion than to switch to opinion 1, as was the case for the PR method. Instead, they have an equal chance of keeping their opinion as to change to the other opinion. This is because both the REF and REC filtering algorithms don't discriminate between the two different opinions of the friends of the nodes with opinion 0. Both opinion 0 and opinion 1 have an equal chance of being promoted on the time-line of the nodes with opinion 0. This reduced tendency of the nodes with opinion 0 to not change their opinion (compared to using the PR method) hampers the ability of the triangle to form an echo chamber. This would explain why, in case of no stubborn nodes, the REC/REF method shows a lower formation of echo chambers compared to the PR method. Now, lets again assume that we make the node with opinion 1 stubborn. In case of the PR filtering algorithm this reduced the formation of echo chambers in the triangle. For the REC/REF method, things are different. The stubborn node with opinion 1 will obviously not change opinion. The two nodes with opinion 0, on the other hand, will now have a higher chance of switching to opinion 1 (which was not the case when using the PR filtering algorithm). This can be seen as follows. The REF/REC filtering algorithms do not discriminate between the two different opinions of the two friends of the nodes with opinion 0. If opinion 1 is, however, stubborn it will keep on pushing that opinion on the time-lines of the two nodes with opinion 0. This will eventually lead to an increase of posts with opinion 1, compared to posts with opinion 0, on the time-lines of the two nodes with opinion 0. The result will then be that the nodes with opinion 0 will have a higher chance of switching to opinion 1, which enhances the formation of echo chambers in the triangle. This would explain the increase in the formation of echo chambers for a small fraction of stubborn people when using the REF or REC filtering algorithms. \\
If too many nodes become stubborn, this increased ability of the system to form echo chambers, in case of using the REF/REC filtering algorithm, starts to decrease again which can also be easily seen with the help of the triangle example. If the fraction of stubborn people becomes too high, we might end up with the case where both the node with opinion 1 and one of the nodes with opinion 0 are stubborn. This strongly reduces the ability to form echo chambers in the triangle since we can no longer have a final situation in which all the three nodes in the triangle carry the same opinion.\\
% make figure 
So far the results for the formation of echo chambers in the system for an initial $50/50$ opinion distribution and introducing a fraction of completely stubborn agents. But what about the evolution of the opinions over time? Are these affected by introducing stubborn people in the system? \\
In Figure \ref{ev_op_50_50_stubb_REC} the evolution of the fraction of nodes with opinion 0 for an initial $50/50$ opinion distribution is shown. Data points are shown together with the standard deviation $\sigma$. This figure shows the results for the REC filtering algorithm. The four network models (ER, WS, SBM (high and low modularity) and SBM-WS) are compared to each other. For each model we look at three different fractions of completely stubborn people: no stubborn people, fracRes $=0$; a fraction of completely stubborn people equal to fracRes $=0.4$ and a fraction of completely stubborn people equal to fracRes $=0.8$. Introducing a fraction of stubborn people is not able to break the initial $50/50$ opinion distribution for neither of the network models.\\

@@

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_PR_8x7.png}
    \caption{ER}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_PR_8x7.png}
    \caption{WS}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_PR_8x7.png}
    \caption{SBM}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_PR_8x7.png}
    \caption{SBM-WS}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$). Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$. Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
\label{ev_op_50_50_stubb_PR}
\end{figure}
\noindent
Figure \ref{ev_op_50_50_stubb_PR} shows the results for the evolution of the fraction of opinion 0 in the system for an initial $50/50$ opinion distribution and when using the PR filtering algorithm. Data points are again represented together with the standard deviation $\sigma$. Three different fractions of stubborn people (fracRes $=0$, fracRes $=0.4$ and fracRes $=0.8$) are compared to each other for the four different network models. The results don't differ from those obtained for the REC filtering algorithm. The status-quo is maintained, the initial $50/50$ opinion distribution does not change over time when introducing stubborn people and using the PR filtering algorithm.\\
\newline
Up till now, we have discussed the results for the formation of echo chambers and the evolution of the opinions when introducing a fraction of completely stubborn nodes in the system for an initial $50/50$ opinion distribution. Lets now look at the results for introducing a fraction of stubborn agents when we are dealing with an unbalanced initial opinion distribution, eg. an initial $20/80$ distribution.\\
\newline

@@

Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80} shows the formation of echo chambers versus the fraction of completely stubborn nodes ($r_i = 1$), fracRes, for an initial $20/80$ opinion distribution (opinion 0 is the minority opinion, opinion 1 is the majority opinion). Figure \ref{PR_frac_compl_stubb_20-80} shows the results for the PR filtering algorithm and Figure \ref{REC_frac_compl_stubb_20-80} shows those for the REC filtering algorithm. The data points are shown together with their standard error $SE$, note that these are very small and almost not visible in the figure. The interpretation of the $y$-axis is slightly different than before. In previous figures, the echo chamber size was interpreted as the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1. Thus we first normalized the echo chambers around opinion 0 and those around opinion 1 and than took the average of both. In case of a $20/80$ distribution we have a very skewed echo chamber distribution, as can be seen in Figure \ref{echo_20_80}, with normalized values going to infinity. We still want to have some average of the echo chambers around opinion 0 and those around opinion 1 to quantify the total formation of echo chambers in the system, but we want to avoid problems with those infinities. Hence, we opted for a slightly different average: first we calculate the total number of like-minded neighborhoods at $t=500$ (thus the sum of the number of echo chambers around opinion 0 and the number of echo chambers around opinion 1) and then we divide that quantity by the total number of like-minded neighborhoods at $t=0$ (hence, we first take the average and than normalize it, instead of normalizing first and then taking the average, as was done before). Note that by taking the average, we neglect the skewed echo chamber distribution. This is, however, not really a problem, since we just want a sense of the total number of echo chambers in the system and how this total number is impacted by the presence of stubborn people and are less interested in the distribution of those echo chambers. \\

@@

When using the PR filtering algorithm, see Figure \ref{PR_frac_compl_stubb_20-80}, we can observe a decrease in the formation of echo chambers with an increasing fraction of stubborn nodes. This behavior is similar to the case of an initial $50/50$ opinion distribution in combination with the PR filtering algorithm, as depicted in Figure \ref{echo_vs_frac_complStubb}. Note that for an initial $20/80$ distribution, in combination with the PR filtering algorithm, the formation of echo chambers in the Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity is non-zero (for a fraction of completely stubborn nodes equal to fracRes $=0$), as can also be observed in Figure \ref{echo_20_80}. This behavior is a consequence of the fact that for an initial $20/80$ opinion distribution, in combination with the PR filtering algorithm, the prevalence of the majority opinion strongly increases, as was already explained for Figure \ref{echo_20_80}. \\
The formation of echo chambers for networks with a high clustering and/or a high modularity is slightly higher. It is however not recommended to compare the formation of echo chambers between the different network models when depicting the results as the average of the echo chambers around opinion 0 and those around opinion 1 in case of an initial $20/80$ opinion distribution. The reason for this is that this average neglects the skewed echo chamber distributions as depicted in Figure \ref{echo_20_80}. When comparing the different network models, these skewed distributions should not be overlooked. The goal of this figure is to show the dependency of the formation of echo chambers in the network models on the presence of a fraction of stubborn people and not to compare the different models against each other. \\

@@

So far the formation of echo chambers in case of an unbalanced initial opinion distribution and introducing a fraction of stubborn people in the system. Lets now discuss the evolution of the opinions in the system. In Figure \ref{ev_op_20_80}, we already saw that for a system with no stubborn agents the REC method does not break the prevalence of the initial opinions in case of an initial $20/80$ opinion distribution. The PR filtering algorithm, on the other hand, resulted in a strong decrease of the prevalence of the minority opinion. Are these results affected by introducing stubborn agents?\\

@@

Figure \ref{ev_op_20_80_stubb_REC} shows the evolution of the fraction of nodes with opinion 0 over time for an initial $20/80$ distribution and when using the REC filtering algorithm. Data point are shown together with their standard deviation $\sigma$. Four network models are compared (ER, WS, SBM (high and low modularity) and SBM-WS). For each network model we compare three different fractions of completely stubborn nodes (fracRes $=0$, fracRes $=0.4$ and fracRes $=0.8$). \\
From this figure it becomes clear that introducing stubborn people in combination with the REC filtering algorithm is not able to break the prevalence of the initial opinions in case of an initial $20/80$ opinion distribution. None of the network topologies, nor the stubborn agents can break the status-quo. This result is similar to the case of an initial $50/50$ opinion distribution in combination with a fraction of stubborn people, as discussed for Figure \ref{ev_op_50_50_stubb_REC}.\\

@@

Figure \ref{ev_op_20_80_stubb_PR} shows the results for the fraction of opinion 0 versus time in case of an unbalanced initial opinion distribution and when using the PR filtering algorithm. Data points are presented along with their standard deviation $\sigma$. Again, we compare four different network models (ER, WS, SBM (high and low modularity) and SBM-WS) and for each model we compare the results for three different fractions of stubborn people (fracRes $=0$, fracRes $=0.4$ and fracRes $=0.8$). \\
If there are no stubborn people, the PR filtering algorithm leads to a significant decrease in the prevalence of the minority opinion in case of an initial unbalanced opinion distribution. Introducing a fraction of stubborn agents changes this scenario. For an increasing fraction of stubborn people, the decrease of the minority opinion becomes less and less strong. This result is the same for all the different network topologies. If there are no stubborn people, the PR method forces nodes with the minority opinion to switch to the majority opinion. If we however introduce a fraction of stubborn people in the system, then we will also end up with some stubborn nodes that carry the minority opinion. These stubborn nodes with the minority opinion will not switch to the majority opinion, no matter how hard the PR filtering algorithm pushes them. The consequence is that these stubborn people with the minority opinion hamper the decrease of that minority opinion.

@@

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_PR_20-80_8x7.png}
    \caption{ER}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_PR_20-80_8x7.png}
    \caption{WS}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_PR_20-80_8x7.png}
    \caption{SBM}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_PR_20-80_8x7.png}
    \caption{SBM-WS}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$). Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$. Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
\label{ev_op_20_80_stubb_PR}
\end{figure}

\newpage
\noindent
\textbf{Making all nodes partly stubborn}\\
\newline
Instead of introducing a fraction of completely stubborn people in the system, as was done before, we will now make all the nodes partly stubborn (so they still have some chance of changing their opinion, they are however less eager to do so). All the nodes are given the same stubbornness parameter $r$ and we let this stubbornness parameter vary from $r = 0$ (all the nodes are completely willing to change opinions) to $r=1$ (all the nodes are completely stubborn).\\
We will first look at the formation of echo chambers and the evolution of the opinions in the system for an initial $50/50$ opinion distribution, where after we will investigate the results for an unbalanced initial opinion distribution ($20/80$).
